2.3 Prob Model for PCR Positivity Estimation

### 2024-05-08 2_scripts/2.3 Prob Model for PCR pos.qmd
Update code for sensitivity analysis and figs.

# lm05r: Abbott--> Roche, using matched CBS
lm05r<-glm(log_roche_n~log_abbott_n , 
          family = "gaussian",
          data = CBSs2)
summary(lm05r)
# model 5r should be focused
```
#### Adjusting the weights of those with negative test results
```{r}
# add a indicator for positivity of two assays
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n > 1]<-"pp"
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n < 1]<-"np"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n < 1]<-"nn"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n > 1]<-"pn"

@@

>>>>

# lm05r: Abbott--> Roche, using matched CBS
lm05r<-glm(log_roche_n~log_abbott_n , 
          family = "gaussian",
          data = CBSs2)
summary(lm05r)
# model 5r should be focused
```

#### Adjusting the weights of those with negative test results
```{r}
# add a indicator for positivity of two assays
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n > 1]<-"pp"
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n < 1]<-"np"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n < 1]<-"nn"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n > 1]<-"pn"

@@

#### 4.2.3 Add regression lines to the figure
```{r}
# Modify p04 for sensitivity analysis
p04<-dta_all %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n),] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study)), alpha=0.5) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="darkgrey") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="darkgrey") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("Jamaica (-)", "Osaka, JP (-)",
                                 "Osaka, JP (+)","UK (+)"), 
                      values = c("red", "blue", "darkgreen", "purple")) +
  theme_bw() + 
  theme(legend.position = "bottom")

p04.2a<-p04 + geom_abline(aes(intercept = -1.60, slope = 0.68, color="Roche->Abbott",),
                  size=0.8) +
      geom_abline(aes(intercept = -1.68, slope = 0.78, color="Abbott->Roche",), 
                  size=0.8) +
scale_colour_manual(values = c("red", "blue", "darkgreen", "purple","blue","red"),
                    guide = guide_legend(nrow = 3,
                            override.aes = list(
                            linetype = c(rep("blank", 4), "solid", "solid"),
                            shape = c(rep(16, 4), NA, NA)))) +
     ggtitle("Jamaica-Osaka-UK matched data")
```
```{r}
# Calculating Beta estimates for model 01r-reverted
(-1)*lm01r$coefficients[1]/lm01r$coefficients[2] #beta_0
1/lm01r$coefficients[2]                          #beta_1

# Calculating Beta estimates for model 05r-reverted
(-1)*lm05r$coefficients[1]/lm05r$coefficients[2] #beta_0
1/lm05r$coefficients[2]                          #beta_1

# Calculating Beta estimates for model 05rw-reverted
(-1)*lm05r.w$coefficients[1]/lm05r.w$coefficients[2] #beta_0
1/lm05r.w$coefficients[2]                            #beta_1
```
```{r}
# Refit models for CBS data, modify p05
p04.2b<-dta_CBS %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n) &
    .$study=="CBS matched",] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study)), alpha=0.40) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="darkgrey") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="darkgrey") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("CBS matched"), 
                      values = c("lightblue")) +
  theme_bw() + 
  theme(legend.position = "bottom")

p04.2b<-p04.2b + geom_abline(aes(intercept = -1.87, slope = 0.64, color="Roche->Abbott",),
                  size=0.8) +
         geom_abline(aes(intercept = 2.02, slope = 2.26, color="Abbott->Roche",), 
                  size=0.8) +
         geom_abline(aes(intercept = -2.14, slope = 0.85, 
                         color="Abbott->Roche (down-weighted)",), 
                  size=0.8, linetype=2) +
         scale_colour_manual(values = c("lightblue","blue","red","red"),
                             guide = guide_legend(nrow=2,
                             override.aes = list(
                             linetype = c(rep("blank", 1), "solid", "solid","dashed"),
                             shape = c(rep(16, 1), NA, NA, NA))
                             )) +
         ggtitle("CBS matched data")
p04.2b
```

```{r}
## make two into one paneled fig
png(file="../4_output/figs/Fig.S1 Sensitivity Analysis.png", 
    width=2400, height=1440, res=300)
cowplot::plot_grid(p04.2a, 
                   p04.2b  + ylab(""), nrow=1,
                   align = 'h',
                   labels = c("A", "B"))
dev.off()
```
